---
title: "514 Lab 4"
author: "Group 13"
date: "Due Date 11/10"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,comment = '', fig.width = 3, fig.height = 3
  )
```


##  Packages

```{r message=FALSE}
# add packages you need for this assignment
install.packages("ggplot2")
library("ggplot2")
```


----- 


## How to work with this document

There are several ways to format your answer, but whatever you do, please make sure it is *readable* by a human and clean. That is, don't leave stray comments and commented instructions in your submitted work. You may use a quote code `>` to start your answer after any r/python code chunks you are typing. An example is provided in the first question below. 

----- 

## Part One: Hypothesis Testing (20 points)

### Checking out some small data sets that come with R 

In this problem, you will load and work with the `mtcars` data set in R.  

Two data samples are independent if they come from unrelated populations and the samples does not affect each other. Here, we assume that the data populations follow the *normal distribution*.
In the data frame column `mpg` (which stands for "miles per galon") of the data set `mtcars`, there are gas mileage data of various 1974 U.S. automobiles. Let's take a look: 

```{r}
 mtcars$mpg 
```

Meanwhile, another data column in `mtcars`, named `am`, indicates the transmission type of the automobile model (0 = automatic, 1 = manual):

```{r}
mtcars$am 
```
In particular, the gas mileage for manual and automatic transmissions are two independent data populations.

Assume that the data in `mtcars` follows the normal distribution, let us look for whether the **difference between the mean gas mileage** of manual and automatic transmissions seems to be statistically significant.

#### a. Please construct a hypothesis test for ratio of the variances of these two populations. Then choose the appropriate test to make a conclusion. (10 points)

> H_0:\sigma^2_{auto}/\sigma^2_{manual} & = 1, \\
H_1:\sigma^2_{auto}/\sigma^2_{manual} & \ne 1,
```{r}
var.test(mpgAuto,mpgManual,alternative = c("two.sided"),conf.level = 0.95)
var.test(mpgAuto,mpgManual,alternative = c("two.sided"),conf.level = 0.9)
```
> If we consider $\alpha = 0.05$, we can not reject $H_0$. However, if $\alpha = 0.1$, based on the result, we should consider to reject $H_0$.


#### b. Based on the result in a), construct a hypothesis test for the means of these two populations. Show your conclusion.(10 points)

> H_0: \mu_{auto}-\mu_{manual} & = 0, \\
H_1: \mu_{auto}-\mu_{manual} & \ne 0,
If we consider $\alpha = 0.05$, we have the variances are the same. Thus we use the condition variances are equal but unknown.
```{r}
t.test(mpgAuto,mpgManual,alternative = c("two.sided"), var.equal = TRUE, conf.level = 0.95)
```
> However, if we choose $\alpha = 0.1$, we should use the condition variances are not equal.
```{r}
t.test(mpgAuto,mpgManual,alternative = c("two.sided"), var.equal = FALSE, conf.level = 0.9)
```
> No matter what value of $\alpha$, based on the result, we should reject $H_0$.

**Hints and shortcuts** 
The gas mileage for automatic transmission can be listed as follows:

```{r, fig.width = 2.5, fig.height = 2.5}
L = mtcars$am == 0 
 mpgAuto = mtcars[L,]$mpg 
 mpgAuto                    # automatic transmission mileage
```

By applying the negation of L, we can find the gas mileage for manual transmission: 

```{r}
mpgManual = mtcars[!L,]$mpg 
 mpgManual                  # manual transmission mileage 
```

Now you should be able to finish solving the problem. Enjoy! :) 



## Part Two: Working With Data (80 points)

In this part, we work on the cybersecurity breach report data downloaded 2015-02-26 from the US Health and Human Services. 

To understand what the data represents, here is some information from the *Office for Civil Rights* of the  *U.S. Department of Health and Human Services*: 

* "As required by section 13402(e)(4) of the HITECH Act, the Secretary must post a list of breaches of unsecured protected health information affecting 500 or more individuals. 
* "Since October 2009 organizations in the U.S. that store data on human health are required to report any incident that compromises the confidentiality of 500 or more patients / human subjects (45 C.F.R. 164.408). These reports are publicly available. Our data  set  was downloaded from the Office for Civil Rights of the U.S. Department of Health and Human Services, 2015-02-26."  

Load this data set and save it as `cyberData`,  using the following code: 

```{r}
cyberData<-read.csv(url("https://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/HHSCyberSecurityBreaches.csv"))
```

### Data Exploration

### Question 1. (5 points)
Check the structure of the data using the `str` command. What type of object is `cyberData`?  How many observations are recorded?
How  many variables are recorded? 
List all  of  the types of random variables that are  recorded based on the output (i.e. int/float etc.). 

```{r}
str(cyberData)
```

> The data set is stored as a data.frame with 1151 rows and 10 columns. There are  1151 observations of 10  variables.  All  of  the types of random variables are: int; chr; logi
  


### Question 2. (20 points)

Let us compare the number of affected individuals across some states. 

* Extract the subset of the data for Kansas and Arkansas; in other words, the subset of the data for which `State` column equals `"KS"` or`"AR"`. Add a third state to the dataframe, say, Illinois (i.e., where `State == "IL"`). Name the new **dataframe** `threeStates`. 


```{r}
cyberDatainARorKS <- cyberData[cyberData$State == "AR" |
cyberData$State == "KS", ]
threeStates <- cyberData[cyberData$State == "AR" | cyberData$State == "KS" |
cyberData$State == "IL", ]
```

* Create a boxplot of `Individuals.Affected` split across the three states. What conclusion can you draw from it?

```{r}
boxplot(cyberData$Individuals.Affected, cyberData$threeStates)
```


The above plot should leave you wondering if Illinois is special, in that it contains some really large data breaches. Let's investigate: 

* How many observations in `cyberData` represent a cyber security breach that affected 100,000 individuals or more? 

```{r}
securityBreachOver100 <- cyberData[cyberData$Individuals.Affected >= 100000, ]
str(securityBreachOver100)
```
> There are 40 observations in cyberData that represent breaches that affected over 100,000 individuals

* How many of those are in Illinois? 


```{r}
securityBreachOver100inIL <- securityBreachOver100[securityBreachOver100$State == "IL", ]
str(securityBreachOver100inIL)
```
> Out of the 40 breaches that affected over 100,000 individuals, 3 of them were in Illinois



### Small analyses across time 

Let us now compare attacks before and after 2013. The goal is to see if there is a significant difference in mean number of affected individuals. 

### Question 3. (10 points)

Check the type of the `Breach.Submission.Date` column: is it a numeric? What type is it? 
```{r}
sapply(cyberData, class)
```
Let us change it to a numeric and extract *the year only*. The code that does this is `as.numeric(format(as.Date(.....),"%Y"))`. Let us use this code to break up the data to before and after 2013, like this: 

```{r}
before2013 <- subset(cyberData, as.numeric(format(as.Date(Breach.Submission.Date),"%Y")) <=2013 )
after2013  <- subset(cyberData, as.numeric(format(as.Date(Breach.Submission.Date),"%Y")) > 2013 )
nrow(before2013)
nrow(after2013)
```


How many observations are in each subset of the population? 

> There are 848 observations before 2013 and 303 observations after 2013



### Specific type of security breaches 

### Question 4. (20 points)

* What proportion of data entries in `cyberData` have `Type.of.Breach  == "Hacking/IT Incident"` ? 

```{r}
numHackIT <-nrow(cyberData[cyberData$Type.of.Breach  == "Hacking/IT Incident",])
numRows <- nrow(cyberData)
noHackIT/numRows
```

> Around 6.69% of the data entries in cyberData are of type: "Hacking/IT Incident"

* What proportion of data entries in `before2013` have `Type.of.Breach  == "Hacking/IT Incident"` ?  

```{r}
numHackITpre2013 <-nrow(before2013[before2013$Type.of.Breach  == "Hacking/IT Incident",])
numHackITpre2013/numRows
```

> Around 4.26% of breached before 2013 are of type: "hacking/IT Incident"

* What proportion of data entries in `after2013` have `Type.of.Breach  == "Hacking/IT Incident"` ?  

```{r}
numHackITafter2013 <-nrow(after2013[after2013$Type.of.Breach  == "Hacking/IT Incident",])
numHackITafter2013/numRows
```

> Around 2.43% of breaches after 2013 are of type: "hacking/IT Incident"


* What are all the different values of `Type.Of.Breach` reported in the data set? How many are hacking/IT incidents? 

```{r}
table(cyberData$Type.of.Breach)
```

> The different type of values for Type.Of.Breach are: "Unknown, Unauthorized Access/Disclosure, Theft, Loss, Improper Disposal, Hacking/IT Incident, Other". There appears to be 77 explicit Hacking/IT Incident and 17 aditional ones were Hacking/IT Incident appears to be one of the types among others. 94 total

* What type of breach is reported in the 748th row of `cyberData`?  How about 349th row? Was row 349 counted in the proportion of Hacking/IT incident breaches you computed above? Why or why not? 

```{r}
cyberData[748,7]
cyberData[349,7]
```

> Row 748: "Loss, Theft" && Row 349: "Hacking/IT Incident, Unauthorized Access/Disclosure". And NO, row 349 was not counted in the proportion computed above, because the calculation was counting entries that ONLY had "Hacking/IT Incident" and rejected those with multiple types.



### Conducting Hypothesis Testing

As you know, this data set contains *all* reports regarding health information data breaches from 2009 to 2015. Let's pretend this is just a *sample* from the population of *all data breaches*, related or not to health information. 

### Question 5. (15 points)

* Compare the number of individuals affected by data breaches (column `Individuals.Affected`) in two states, Arkansas (`State=="AR"`) and California (`State=="CA"`). 
This can be done by performing a test of difference in means, for example. Assume the individuals affected follows an **approximately normal distribution**.


*Please note, in order to answer this question completely, you will need to run several lines of code, extract subsets of the data appropriately, run a statistical hypothesis test, and interpret the results. Draw a conclusion. Partial answers to the question are insufficient.* 

```{r}
# Your code here.
# 
```

> Your answer / discussion / interpretation here. 
