---
title: "514 Lab 1"
author: "Joseph Fulkerson, Julio Pagan"
date: "Due Date 9/13"
output: 
        html_document: default
        pdf_document: default
---

```{r message=FALSE}
# If you don't already have the tidyverse library installed,
# you will need to type install.packages("tidyverse") into the Console
library(tidyverse) 
```

### 1. Changing the author field and file name. (5 points)

##### (a) Change the `author:` field on the Rmd document 

##### (b) Rename this file to "HW1_YourGroupNumberHere.Rmd", where YourGroupNumberHere is changed to your group number (e.g. Group1).

### 2.  Hello World! (5 points)

Here's an R code chunk that prints the text 'Hello world!'.

```{r}
print("Hello world!")
```

#### (a) Modify the code chunk below to print your name
```{r}
print("Julio Pagan, Joseph Fulkerson")
```

<br>

### 3.  Creating a numeric vector (30 points)
We just learned about the `c()` operator, which forms a vector from its arguments.  If we're trying to build a vector containing a sequence of numbers, there are several useful functions at our disposal.  These are the colon operator `:` and the sequence function `seq()`.

##### `:` Colon operator: 
```{r}
1:10 # Numbers 1 to 10
127:132 # Numbers 127 to 132
```

##### `seq` function: `seq(from, to, by)`
```{r}
seq(1,10,1) # Numbers 1 to 10
seq(1,10,2) # Odd numbers from 1 to 10
seq(2,10,2) # Even numbers from 2 to 10
```

> To learn more about a function, type `?functionname` into your console.  E.g., `?seq` pulls up a Help file with the R documentation for the `seq` function.

#### (a) Use `:` to output the sequence of numbers from 3 to 12
```{r}
3:12
```

#### (b) Use `seq()` to output the sequence of numbers from 3 to 30 in increments of 3
```{r}
print(seq(3,30,3))
```

#### (c) Save the sequence from (a) as a variable `x`, and the sequence from (b) as a variable `y`.  Output their product `x*y`
```{r}
x <- 3:12
y <- seq(3,30,3)
print(x*y)
```

<br>

### 4. Cars data (60 points)

We'll look at data frame and plotting in much more detail in later classes.  For a previous of what's to come, here's a very basic example.

For this example we'll use a very simple dataset.  The `cars` data comes with the default installation of R.  To see the first few columns of the data, just type `head(cars)`.

```{r}
head(cars)
```

#### (a) Calculate the average and standard deviation of `speed` 

```{r}
speed_average <- mean(cars$speed)
speed_standard_deviation <- sd(cars$speed)
speed_results <- print(paste("The average of speed is " , speed_average , " and the standard deviation is " , speed_standard_deviation))
```

> The average of speed is 15.4 and the standard deviation is 5.28764443523478

# @jpagan2 ______________________________________________________________
<br> 

#### (b) Calculate the average and standard deviation of `dist` 

```{r}
average_dist <- mean(cars$dist)
standard_dev_dist <- sd(cars$dist)
answer_dist <- print(paste("The average of dist is " , average_dist , " and the standard deviation is " , standard_dev_dist))
```

> The average of dist is 42.98 and the standard deviation is 25.7693774920259

<br> 

#### (c) Calulate the average and standard deviation of `dist` when `speed`  is greater than the average. Compare the results with the answers in (b)

```{r}
# Edit me
```

> Type your answer here

<br> 

We can easily produce a histogram of stopping distance using the `qplot` function (built-in `tidyverse` package).

```{r}
qplot(cars$dist) # Histogram of stopping distance
```

How to change the number of bins?

```{r}
qplot(cars$dist,bins = 20)
```


#### (d) Produce a histogram of speed using the `qplot` function with **5 bins**.

```{r}
qplot(cars$speed,bins = 5)
```

The `qplot(x,y,...)` function can also be used to plot a vector `y` against a vector `x`.  You can type `?qplot` into the Console to learn more about the basic qplot function. 

#### (e) Use the `qplot(x,y)` function to create a scatterplot of dist against speed.

```{r}
qplot(cars$dist,cars$speed)
```

#### (f) Use the `boxplot` function to create a boxplot of speed.

```{r}
boxplot(cars$speed)
```

